---
title: "Avetis Mishegyan Final Independent Project"
author: "Avetis Mishegyan"
date: '`r Sys.Date()`'
output:
  rmdformats::readthedown:
    toc_depth: 3
    self_contained: TRUE
    thumbnails: TRUE
    lightbox: TRUE
    gallery: TRUE
    highlight: espresso
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fig.path = "../output/")
```

## Introduction

Today we are going to present our final independent projects for Biol551 which should have:

* A new stand-alone public repository on the BIOL551 organization on GitHub
* At least 3 outputs associated with your project
  * it can be a plot, table, animation, unique useful functions, etc.
* Give a 10-minute presentation to the class explaining the data set used  

The data set that I will be using is based off my research project. My project's purpose is in the investigation of the genetic and physiological changes in flood-stressed turmeric plants as compared to control turmeric. The following data consists of the 3 physiological assays using turmeric leaf samples. These assays include soluble sugar content, chlorophyll content, and catalase activity.


## Load Libraries

Here is where we will load any libraries that we will be using. These libraries contain use full packages that are collections of R functions, data, and compiled code in a well-defined format.

```{r}
library(tidyverse)
library(here)
library(performance) # for checking model assumptions
library(alphonse) # for alphonse color palette
library(viridis) # for viridis color palette
```

## Load Data

Here is where we will load (ie. pull) our data as a ".csv" format into our R script. Additionally, this is where we would also inspect our data.

```{r}
dil_chloro_data <- read_csv(here("avetis_project", "data", "diluted_turmeric_flood_chlorophyll_test_2022_12_23.csv")) # reads in data from a .csv file
glimpse(dil_chloro_data) # allows us to inspect the data

catalase_data <- read_csv(here("avetis_project", "data", "turmeric_flood_catalase_test_2022_12_23.csv"))
glimpse(catalase_data)

sugar_data <- read_csv(here("avetis_project", "data", "turmeric_flood_sugar_test_2022_12_23.csv"))
glimpse(sugar_data)
```

## Data Analysis

### Chlorophyll Content Assay

#### Total Chlorophyll Function

```{r}
total_chlorophyll <- function(abs_645, abs_663) { # name of function, and list of inputs (ie. arguments)
  chloro_a <- 12.7*{{abs_663}} - 2.69*{{abs_645}} # code you have developed in body of the function
  chloro_b <- 22.9*{{abs_645}} - 4.68*{{abs_663}}
  total <- chloro_a + chloro_b
  return(total) # returns values we want
}

total_chlorophyll(.2, .8) # test of function
```

#### Total Chlorophyll Calculations

```{r}
mean_total_calc_dil <- dil_chloro_data %>%
  pivot_wider(names_from = abs,
              values_from = chloro_dil) %>%
  mutate(total_dil = total_chlorophyll(abs_645, abs_663)*5*10) %>%
  group_by(treatment) %>%
  summarise(mean_total_dil = mean(total_dil, na.rm = TRUE),
            sd_total_dil = sd(total_dil, na.rm = TRUE))
glimpse(mean_total_calc_dil)
```

#### Bar Plot Function

```{r}
bar_plot <- function(data, x, y, fill, sd, lab_x, lab_y, title){ 
ggplot(data, aes(x = {{x}},
                 y = {{y}}, 
                 fill = {{fill}})) +
  geom_bar(stat = "identity") + # bar plot function
  geom_errorbar(aes(ymin = {{y}} - {{sd}}, # error bar function
                    ymax = {{y}} + {{sd}}),
                width = .2) + # adjusts width of error bars
  labs(x = lab_x,
       y = lab_y,
       title = title) +
  scale_fill_manual(values = alphonse("demon")) + # colors bars using alphonse color palette
  theme_classic() + # sets plot theme to classic
  theme(legend.position = "none") + # removes legend
  theme(plot.title = element_text(hjust = .5)) # centers plot title
}
```

#### Total Chlorophyll Bar Plot

```{r total_chloro_bar, fig.width=10, fig.height=5, fig.align='center',  fig.cap="Figure 1. The mean total chlorophyll of extractions of all biological replicates consisting of 3 separate leaf samples for each biological replicate for each treatment is represented above. Biological replicates consisted of control: n = 3 and flood: n = 3. Error bars above represent standard deviation."}
bar_plot(data = mean_total_calc_dil,
         x = treatment,
         y = mean_total_dil,
         fill = treatment,
         sd = sd_total_dil,
         lab_x = "Treatment",
         lab_y = "Mean Total Chlorophyll (mg)",
         title = "Turmeric Mean Total Chlorophyll Flood v. Control")
```

#### Total Chlorophyll Significance Analysis

```{r total_chloro_mod_check, fig.width=10, fig.height=10, fig.align='center'}
raw_total_calc_dil <- dil_chloro_data %>%
  pivot_wider(names_from = abs,
              values_from = chloro_dil) %>%
  mutate(total_dil = total_chlorophyll(abs_645, abs_663)*5*10)
glimpse(raw_total_calc_dil)

total_chloro_mod <- lm(total_dil ~ treatment, data = raw_total_calc_dil) # runs a simple linear model

check_model(total_chloro_mod) # checks assumptions of an lm model

t.test(total_dil ~ treatment, data = raw_total_calc_dil) # runs a t-test analysis
```
### Sugar Content Assay

#### Sugar Content Violin and Boxplot

```{r sugar_violin_box, fig.width=10, fig.height=5, fig.align='center',  fig.cap="Figure 2. The median absorbance of soluble sugar of all biological replicates consisting of 3 separate leaf samples for each biological replicate for each treatment is represented above. Biological replicates consisted of control: n = 3 and flood: n = 3."}
sugar_data %>% 
  ggplot(aes(x = treatment,
             y = sugar,
             fill = treatment)) +
  geom_violin(width = 1.4) + # violin plot function and adjusts width of box plot
  geom_boxplot(width = 0.1, # box plot function and adjusts width of box plot
               color = "grey", # sets color of box plot
               alpha = 0.2) + # sets opacity of box plot
  labs(x = "Treatment",
       y = "Absorbance (630 nm)",
       title = "Turmeric Soluable Sugar Content Flood v. Control") +
  scale_fill_viridis(discrete = TRUE) + # colors violin using viridis color palette
  theme_classic() +
  theme(legend.position = "none") +
  theme(plot.title = element_text(hjust = .5))
```

#### Sugar Significance Analysis

```{r sugar_data_mod_check, fig.width=10, fig.height=10, fig.align='center'}
sugar_data_mod <- lm(sugar ~ treatment, data = sugar_data)

check_model(sugar_data_mod)

t.test(sugar ~ treatment, data = sugar_data)
```
### Catalase Acivity Assay

```{r catalase_box_jitter, fig.width=10, fig.height=5, fig.align='center',  fig.cap="Figure 3. The median effervescence volume of all biological replicates consists of biological replicates of control: n = 3 and flood: n = 3."}
catalase_data %>%
  ggplot(aes(x = treatment,
             y = efferves,
             fill = treatment)) +
  geom_boxplot() +
  geom_jitter(color = "black", # jitter plot function and sets color of jitter points
              size = 1, # sets size of jitter points
              alpha = 1) + # sets opacity of jitter points
  labs(x = "Treatment",
       y = "Effervescence Volume (mL)",
       title = "Turmeric Catalase Activity Flood v. Control") +
  scale_fill_manual(values = alphonse("jojo")) + # colors bars using alphonse color palette
  theme_classic() +
  theme(legend.position = "none") +
  theme(plot.title = element_text(hjust = .5))
```

#### Catalase Significance Analysis

```{r catalase_data_mod_check, fig.width=10, fig.height=10, fig.align='center'}
catalase_data_mod <- lm(efferves ~ treatment, data = catalase_data)

check_model(catalase_data_mod)

t.test(efferves ~ treatment, data = catalase_data)
```
